---
title: "lab week 5 time series and a map"
author: "Ruoyu Wang"
date: "2020/2/6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)
```

## monthly US energy consumption (renewables)

```{r}
us_renew <- read_csv(here::here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()
```

- make description all lowercase
- only keep observations where "description" variable contains "consumtion"
- remove any observations where "description" variable contains "total"

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total"))
```


### Convert 'yyyymm' column to a date

```{r}
renew_date <-renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>%  # just pull out year and month, class will be yearmonth and date
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)  # specify the column, otherwise any rows have a na will be deleted

# make a version wher I have the month & year in separate columns
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>%  # label --> write in text not numbers
  mutate(year = year(yr_mo_day))
```

### look at it

```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month_sep, y = value))+  # or aes(group = description)
  geom_line(aes(color = description))+
  theme_minimal()


renew_gg
```

updating colors with paletteer palettes:
```{r}
renew_gg +
  scale_color_paletteer_d("basetheme::brutal")  # package_name::palette name
```

### Coerce renew_parsed to a tsibble
```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep) # index, the time column
```

let's look at our ts data in a couple different ways:

```{r}
renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value) 
renew_ts %>% gg_season(value)

# make the season plot in ggplot
ggplot(data = renew_parsed, aes(x = month, y = value, group = year))+
  geom_line(aes(color = year))+
  facet_wrap(~description,
             ncol = 1,
             scales = "free",
             strip.position = "right") # great! change the facet proporties!
```

### Just look at the hydeoelectric energy consumption
```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)

```

### What if I want quariterly average consumption for hydro?
```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% # great way to group, new_column = ~
  summarize(avg_consumption = mean(value))

head(hydro_quarterly)
```

### Decompose that hydro_ts

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

components(dcmp) %>% autoplot()
```

```{r}
hist(components(dcmp)$remainder)
```

Now lookat the ACF
```{r}
hydro_ts %>% 
  feasts::ACF(value) %>% 
  autoplot()
# 12 months later have the most correlation
```

# DANGER DANGER do more research before making assumption
```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value),
    ETS(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010)) + theme_minimal()
```

### Make a world map!

```{r}
world <- read_sf(dsn = here::here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)
```

